---
title: "Question 3d and 3e"
author: 'Mufitcan Atalay '
date: "February 19, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

demand = read.csv("demand_data.csv")

#multinominal log likelihood

log_l = function(beta){
  ll = 0 
  dataset = demand
  y = dataset[,14]
  x_ij = dataset[,2:10]
  for (i in 1:length(y)){
    second_term = log(1 + exp(as.matrix(x_ij[i, 1:3])%*%beta[1:3]) + exp(as.matrix(x_ij[i, 4:6])%*%beta[4:6]) + exp(as.matrix(x_ij[i, 7:9])%*%beta[7:9]))
    if (y[i] != 0 ){
      x_j = as.matrix(x_ij[i,(3*y[i]-2):(3*y[i])])
      first_term = x_j%*%beta[(3*y[i]-2):(3*y[i])]
      ll = ll + (first_term - second_term) 
    } else {
      ll = ll - second_term
    }
  } 
  return(ll[1,1])
}

#gradient score_function
score_function = function(beta){
  dataset = demand
  y = dataset[,14]
  x_ij = dataset[,2:10]
  score_vector = rep(0,9)
  for (i in 1:length(y)){
    denom = 1 + exp(as.matrix(x_ij[i, 1:3])%*%beta[1:3]) + 
      exp(as.matrix(x_ij[i, 4:6])%*%beta[4:6]) + 
      exp(as.matrix(x_ij[i, 7:9])%*%beta[7:9])
    if (y[i] != 0){
      x_j = as.matrix(x_ij[i,(3*y[i]-2):(3*y[i])])
      numerator = exp(x_j%*%beta[(3*y[i]-2):(3*y[i])])
      for (j in 1:length(beta)){
        if (y[i] == j){
          indicator = 1
        } else {
          indicator = 0
        }
        x_m = x_ij[i,j]
        row_score = x_m*(indicator - numerator/denom)
        score_vector[j] = score_vector[j] + row_score
      } 
    } 
  }
  return(score_vector)
  }

#library(optimx)
#utput = optimx(par = rep(1,9),
 #               fn = log_l,
  #              gr = score_function)

```
When we try to optimize the function gives an error message. There is a small bug in our score function but we couldn't figure it out in many hours of debugging. 

```{r}
#calculating shares
library(dplyr)
shares = table(demand$marketindex, demand$choice)
shares = shares/100
marketindex = seq(1, 50, length.out = 50)
y_1 = data.frame(cbind(marketindex, log(shares[,2]/shares[,1])))
y_1 = y_1 %>% filter(V2 != Inf) %>% filter(V2 != -Inf) %>% filter(!is.na(V2))
  
y_2 = data.frame(cbind(marketindex, log(shares[,3]/shares[,1])))
y_2 = y_2 %>% filter(V2 != Inf) %>% filter(V2 != -Inf) %>% filter(!is.na(V2))

y_3 = data.frame(cbind(marketindex, log(shares[,4]/shares[,1])))
y_3 = y_3 %>% filter(V2 != Inf) %>% filter(V2 != -Inf) %>% filter(!is.na(V2))
    
data_1 = unique(left_join(y_1, demand[,c(1,2:4)], by="marketindex"))
data_2 = unique(left_join(y_2, demand[,c(1,5:7)], by="marketindex"))
data_3 = unique(left_join(y_3, demand[,c(1,8:10)], by="marketindex"))

#OLS for j = 1
j_1 = lm(V2 ~ x1_prod1 + x2_prod1 + x3_prod1, data=data_1) 
summary(j_1)
#j=2
j_2 = lm(V2 ~ x1_prod2 + x2_prod2 + x3_prod2, data=data_2) 
summary(j_2)
#j=3
j_3 = lm(V2 ~ x1_prod3 + x2_prod3 + x3_prod3, data=data_3) 
summary(j_3)
```
The folllowing results for the estimates make sense. The coefficient estimate for price is always negative and the other coefficient estimates for product quality are all positive. This is in line with what we would expect.
