---
title: 'Stat 245: Question 7'
author: "Mufitcan Atalay"
date: "4/18/2018"
output:
  pdf_document: default
  html_document: default
---
_*a)*_
```{r}
(1-pnorm(10, mean = 0, sd = 1))*2
```
This is the tenth standard deviation so it is understandable that we get a p-value that is so small that `R` rounds it 0. 
_*b)*_
```{r}
(1-pnorm(1, mean = 0, sd = 1))*2
```
This makes sense since we are looking at the 1st standard deviation from the assumed mean of $0$.

_*c)*_
```{r}
rv1 <- rnorm(100, mean = 50, sd =1)
mrv1 <- mean(rv1)
z1 = abs((mrv1 - 50)/(1/sqrt(100)))
pval1 <- (1-pnorm(z1, mean = 0, sd = 1))*2
pval1
```


_*d)*_
```{r}
nrSamples = 1000 #number of repetitions
rvs <- lapply(1:nrSamples, function(x) rnorm(n=100, mean = 50, sd = 1 )) # Creating 1000 100-sized samples from our normal distribution
means <- sapply(rvs, mean) # Obtaining 1000 means from our samples
zs = abs((means - 50)/(1/sqrt(100)))
pvals <- (1-pnorm(zs, mean = 0, sd = 1, lower.tail = ))*2 # Obtainig 1000 p-values
hist(pvals, main = "Histogram of P-values from 1000 Simulations", xlab = "P-values", ylab = "Frequency") # Plotting P-values
```

_*e)*_
```{r}
rv2 <- rnorm(100, mean = 0, sd =1)
mrv2 <- mean(rv2)
z2 = abs((mrv2)/(1/sqrt(100)))
pval2 <- (1-pnorm(z2, mean = 0, sd = 1))*2
pval2
```
_*f)*_
```{r}
nrSamples = 1000 #number of repetitions
rvs1 <- lapply(1:nrSamples, function(x) rnorm(n=100, mean = 0, sd = 1 )) # Creating 1000 100-sized samples from our normal distribution
means1 <- sapply(rvs1, mean) # Obtaining 1000 means from our samples
zs1  = abs((means1)/(1/sqrt(100)))
pvals1 <- (1-pnorm(zs1, mean = 0, sd = 1))*2 # Obtainig 1000 p-values
hist(pvals, main = "Histogram of P-values from 1000 Simulations", xlab = "P-values", ylab = "Frequency") # Plotting P-values
```

Finding the number of values below $\alpha=0.05$
```{r}
sum(pvals1 < 0.05)
```
There are rougly $50$ p-values below $0.05$ (when I knit the document this number is randomly generated so it is impossible to really know the value, but it is always close to 50). This is because p-values form a uniform distribution when the null hypothesis is true, so the number of p-values below $0.05$ should equal the probability $\alpha =0.05$ times our size of a $1000$. The whole point of a p-value is that we are able to take a z-test or any other form of test and turn it into a value which we can compare to our $\alpha$, so it must follow a uniform distribution $U(0,1)$. Since our the $\alpha$ we select is $0.05$, there should only be $0.05$ proportion out of all the p-values which are below $0.05$, which is the case. 

_*g)*_
As stated above the point of p-values is that you are able to take any test statistics and compare it to a desired confidence level $\alpha$. This means that p-values must follow a uniform distribution that is bound between $0$ and $1$, which is what we get from our answeres to  _*d)*_ and _*f)*_. In this way the  proportion of p-values below our desired level $\alpha$ will always be roughly aroound $\alpha$. 
